Effects of PCSK9 Inhibition With Evolocumab on Lipoprotein Metabolism in subjects with elevated Lp(a)

### 2019-10-10 manuscript.Rmd
discussion for draft version 3

Evolocumab treatment for 16 weeks resulted in mean 18.9% (95% credible interval: 7.51% to 29.2%) reduction in Lp(a) compared with placebo.

@@

In the current study, evolocumab treatment for 16 weeks resulted in on average 18.9% reduction in Lp(a) compared with placebo.
Our detailed lipoprotein profiling revealed that treatment of evolocumab substantially reduced concentrations of VLDL, IDL and LDL particles. 
Our observation is consistent with the previous study, in which evolocumab treatment not only reduced concentrations of LDL particles, but also substantially reduced triglyceride-rich lipoprotein particles [@Toth2018].
The next question we came up with is "why we observed large effect of evolocumab on VLDL, IDL and LDL particle concentrations but relatively small effect on Lp(a)?" 
To address the question, we analyzed the composition of LDL particles. 
Because Lp(a) is often thought to be comprised of a LDL particle and a covalently bound apo(a), we assumed that a pool of LDL particles contained LDL particles with and without apo(a) tail. 
At baseline, we identified that on average about 50% of large, medium and small LDL particles had apo(a) tail. 
Evolocumab treatment not only decreased concentrations of LDL particles, but also resulted in 10-20% increasing in LDL particles with apo(a). 
The composition shifting to relatively more LDL particles with apo(a) indicated that LDL particles with apo(a) were less preferably cleared from the circulation compared to their counterparts without apo(a). 
Indeed Lp(a) needs to compete with LDL particles for binding to the LDL receptor, and is preferred for uptake by the LDL receptor when the LDL particle concentration is low [@Raal2016].  
We hypothesized that the LDL particles with and without apo(a) tail may have different affinity for binding to the LDL receptor. 
Besides LDL receptor, we acknowledge that VLDL receptor can also mediate Lp(a) catabolism [@Argraves1997]. 
In our analysis, we observed substantial reduction in VLDL particle concentrations after treatment of evolocumab. 
Since PCSK9 can induce the degradation of LDL receptor as well as VLDL receptor, evolocumab treatment must enhance activity of both receptors [@Poirier2008]. 
Interestingly, we observed that most of patients receiving evolocumab ended in very low concentration (below the detection limit) of extremely large and very large VLDL particles, suggesting enhancement of VLDL receptor activity.

@@

Strengths and limitations of our study warrant consideration. 
Our study provided first insight into relationship between Lp(a) concentrations and NMR LDL particle concentrations. 
One limitation of our study is that due to small sample size, our uncertainty of LDL particle composition estimation is large.

>>>>

Evolocumab treatment for 16 weeks resulted in mean 17% (95% credible interval: 8% to 26%) reduction in Lp(a) compared with placebo.

@@

The major finding of this work is that evolocumab treatment in patients with high Lp(a) massively decreases large VLDL in addition to the well known effect on LDL. 
In the current work, evolocumab treatment resulted in on average 17% reduction in plasma Lp(a) concentration compared with placebo. 
Our detailed metabolomic profiling revealed that Lp(a) reduction by PCSK9 inhibition corresponded to substantial reduction in concentrations of VLDL, IDL and LDL particles as well as their lipid and fatty acid contents. 
Our observation is consistent with the previous study, in which evolocumab treatment not only reduced concentrations of LDL particles, but also substantially reduced triglyceride-rich lipoprotein particles [@Toth2018]. 
Since a PCSK9 monoclonal antibody such as evolocumab reduces the degradation of the LDL receptor as well as the LDL receptor, the very low concentration of VLDL particles after the treatment may be ascribed to the increased umber of VLDL receptors [@Poirier2008]. 
The upregulation of VLDL receptors can enhance uptake of VLDLs but also Lp(a) from the circulation [@Argraves1997]. 
In addition, our analysis also showed that treatment of evolocumab lowered concentrations of very large HDL particles. 
Very large HDL is cholesterol enriched and mainly captured by the SR-B1 present in the liver [@Acton1996]. 
Similar to the VLDL receptor, SR-B1 has also been shown to bind Lp(a), and its upregulation will probably enhance the Lp(a) catabolism [@Yang2013].

In our analysis, we observed a large discordance in Lp(a) and LDL reduction in response to evolocumab. 
This discordance was suggested by a recent study to be related to alternative pathways of Lp(a) reduction beyond LDL receptor mediated clearance [@Shapiro2019]. 
Lp(a) and LDL significantly overlap in density, as well as protein and lipid composition. 
As a result, current technologies including the NMR metabolomics used in this study cannot distinguish these two lipoproteins, and include the content of Lp(a) in the LDL measurements. 
We think that it is important to separate the Lp(a) content from the LDL measurements in order to better understand the discordance between these two lipoproteins in response to evolocumab. 
To this end, we developed a statistical model based on the assumption that Lp(a) was comprised of a LDL particle and a covalently bound apo(a). 
At baseline, we identified that on average about 50% of large, medium and small LDL particles had apo(a) tail. 
After evolocumab treatment, on average about 60%-70% of large, medium and small LDL particles having apo(a) tail. 
The relatively more LDL particles with apo(a) indicated that LDL particles with apo(a) were less preferably cleared from the circulation compared to their counterparts without apo(a). 
Indeed a recent study showed that LDL without apo(a) was cleared faster than Lp(a), which was found to compete with LDL particles for binding to the LDL receptor [@Viney2018; @Raal2016].

@@

Strengths and limitations of our study warrant consideration. 
Our study provided the first NMR metabolomics data for a PCSK9 inhibition trial. 
The detailed metabolomic profiling revealed metabolic effects of the PCSK9 inhibitor, and improved our understanding of this lipid-lowering therapy. 
Moreover, in this study we developed a statistical method to separate Lp(a) content from the LDL measurements. 
However, due to the heterogeneity of Lp(a) and limited number of participants, our estimation of the proportions of LDL with apo(a) had very large uncertainty. 
Applying our statistical model with a larger number of participants as well as validation in the laboratory are warranted in the future studies.

In conclusion, our NMR metabolomic profiling revealed that Lp(a) reduction by PCSK9 inhibition corresponded to substantial reduction in VLDL, IDL and LDL particles as well as their lipid contents. 
Our analysis also suggested that LDL with and without apo(a) had different preference to be cleared from the circulation.
